Modeling Human Interactions in Collaborative Interactive Evolutionary Computation

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Addressing JJ comments

Three case studies are presented as a proof of concept, providing both
conceptual and implementation details of the framework. 
The experiments were implemented using the EvoSpace-Interactive framework \cite{garcia2013evospace}, 
and where employed in different contexts of interaction. 
Additional details of each case study application is presented elsewhere refs[][][]. 
Giving this paper the objective of presenting the underlying conceptual framework.

The rest of this paper is organized as follows.
Section \ref{sec:related} presents related work on the topic 
of Interactive Evolution.
Then, Section \ref{sec:framework} presents the framework of Human Interactions in Collaborative Interactive 
 Evolutionary Computation which 
the main proposal of this work. An implementation using the EvoSpace-Interactive  framework is detailed in Section \ref{sec:implementation}.
The case studies  and results are presented in Section \ref{sec:experiments}.
Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.

@@

\begin{figure*}[!t]
    \centering
        \includegraphics[width=4.5in]{img/framework.eps}
    \caption{IEC Human-centered framework.}
    \label{fig:uc_framework}
\end{figure*}

\section{Human-Centered IEC Framework}
\label{sec:framework}
% Comment: 
% Maybe in this case we should use the term software framework or architecture. 
The general goal of this line of research is to develop a human-centered \cite{gasson2003human} 
software framework for interactive evolutionary computation (IEC). 
A framework is defined as a reusable architectural design together
with an implementation \cite{campbell1991choices}, in this case 
providing generalized components to developers of IEC systems. In the
description below %below? If that's te chase, we _will_ - JJ
% Not only in the description, more like a paradigm shitf. But ok for now - Mario
 we will deliberately use the more general 
term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
could just be interacting naturally with their environment and
unwittingly and at the same time they are implicitly
evaluating a possible solution.

%Which framework? Are we talking about a particular framework or that
%is the general rule for frameworks? - JJ
%Is general bu I was talking about this one - Mario
The proposed framework includes components that can be refined to increase
participation and also to minimize the amount of evaluations needed for the evolutionary 
process in a given IEC application. Software frameworks often have a vison or a way 
of thinking \cite{carneiro2010introducing} guiding their
design. Before diving into details the main design decisions of the
framework are explained next:  % constraints? Rules? You don't decide
                               % that users are humans, they simply
                               % are - JJ 
                               % Ok, is more like we want to acknowledge that, 
                               % To Do - Mario

@@

\item {\bf Users are volunteers} Users donate their computing resources, so they are unaccountable, 
  some times they try to game the system. Providers must actively promote and design the interactive
  system to engage volunteers \cite{oh2015clicking}. % Define
                                % providers somewhere. Experiment
                                % owners? - JJ

@@

% Say something about how this is important for the framework and how
% you incorporate it - JJ
The human centered IEC architecture consists of three high level components depicted
in figure \ref{fig:uc_framework}:

@@

\subsection{EvoDraw}
As a case study, a IEA was developed using the 
EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}.
A brief description of the application is presented next, focusing
on the data elements that where ported to the graph model.

\subsection{Fitness Assignment}
\label{sec:assignment}
The ES-I % You should use always the same acronym, not start using it
         % here - JJ
 platform employs a collaborative methodology for performing fitness
 assignment. Several registered users assign a quality assessment to a single
phenotype and then an aggregated fitness value is calculated,
depending on the tastes of each particular user,  % how do you define
                                % "taste"? - JJ
resulting in a many-to-many relationship between users and
phenotypes. %scored by them? - JJ 
Many systems % formerly published methods? Other researchers? - JJ
query this user-phenotype relation to extract relevant
knowledge about the process and the population, for instance showing the
most popular phenotype, or the the user with more participation 
\cite{picbreeder}.
In order to do this, data about phenotypes % metadata better? - JJ
must be permanently stored, even
if they are no longer in the active population.

\subsection{Collaboration}
\label{sec:col}
Users need to authenticate themselves to the system using their
Facebook account. In that sense, some users might not be interested
either because they do not want to give that information or simply because
they do not use that particular social network. Even as we might lose
some users that way, the additional information we obtain for scoring
phenotypes more than compensates for that. % Added this, check it
                                % out. 
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those phenotypes they like, or by taking phenotypes
from their friend's collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, a phenotype sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a quality assessment % Why don't you use other
                                % thing instead of quality? We are not
                                % talking about quality, but about
                                % tastes or similar - JJ
 to the phenotype using
a five star rating system or,
additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special folder that stores those phenotypes a user likes and wishes
to save. After the user finishes interacting with the phenotype
on the Wall, he can choose to retrieve a new one from the population.
At the left hand side of \ref{fig:web}, the web page shows the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, users can browse the content of each collection and from
there share images through their social network.
When a user hovers the cursor over a phenotypes a pop-up pane shows how many users have
liked the phenotype. The pane also includes a link to the phenotype's 
details, the parents, genetic operators that created it, and genealogy
information. This makes the assignment of fitness through the rating
system a {\em social} activity, pursuing the objective of this work,
which is to increase engagement of the users.
\subsubsection{Graph Model}

>>>>

Three case studies are presented as a proof of concept, providing both
conceptual and implementation details of the framework. 
The experiments were implemented using the EvoSpace-Interactive (ES-I) framework \cite{garcia2013evospace}, 
and where employed in different contexts of interaction. 
Additional details of each case study application is presented elsewhere refs[][][]. 
Giving this paper the objective of presenting the underlying conceptual framework.

The rest of this paper is organized as follows.
Section \ref{sec:related} presents related work on the topic 
of Interactive Evolution.
Then, Section \ref{sec:framework} presents the framework of Human Interactions in Collaborative Interactive 
 Evolutionary Computation which 
the main proposal of this work. An implementation using the ES-I framework is detailed in Section \ref{sec:implementation}.
The case studies  and results are presented in Section \ref{sec:experiments}.
Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.

@@

\begin{figure*}[!t]
    \centering
        \includegraphics[width=4.5in]{img/framework.eps}
    \caption{IEC Human-centered framework.}
    \label{fig:hc_framework}
\end{figure*}

\section{Human-Centered IEC Framework}
\label{sec:framework}
% Comment: 
% Maybe in this case we should use the term software framework or architecture. 
The general goal of this line of research is to develop a human-centered \cite{gasson2003human} 
software framework for interactive evolutionary computation (IEC). 
A framework is defined as a reusable architectural design together
with an implementation \cite{campbell1991choices}, in this case 
providing generalized components to developers of IEC systems. 
% Just by saying is Human Centered this is not needed, it was from the User-Centered days - Mario
%In the
%description below %below? If that's te chase, we _will_ - JJ
% Not only in the description, more like a paradigm shitf. But ok for now - Mario
% we will deliberately use the more general 
%term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
%could just be interacting naturally with their environment and
%unwittingly and at the same time they are implicitly
%evaluating a possible solution.

%Which framework? Are we talking about a particular framework or that
%is the general rule for frameworks? - JJ
%Is general bu I was talking about this one - Mario
The proposed framework includes components that can be refined to increase
participation and also to minimize the amount of evaluations needed for the evolutionary 
process in a given IEC application. Software frameworks often have a vison or a way 
of thinking \cite{carneiro2010introducing} guiding their
design. Before diving into details the main design considerations of the
framework are explained next:  % constraints? Rules? You don't decide
                               % that users are humans, they simply
                               % are - JJ 
                               % Ok, is more like we want to acknowledge that, 
                               % considerations? - Mario

@@

\item {\bf Users are volunteers} Users donate their computing resources, so they are unaccountable, 
  some times they try to game the system. Project owners must actively promote and design the interactive
  system to engage volunteers \cite{oh2015clicking}. % Define
                                % providers somewhere. Experiment
                                % owners? - JJ
                                % Using project, like BOINC  https://boinc.berkeley.edu/trac/wiki/VolunteerComputing - M

@@

% Say something about how this is important for the framework and how
% you incorporate it - JJ
% Done - Mario
The above considerations have guided the design of the framework, and througout they have been
treated as application constraints. In order to satisfy theses requirements the
human centered IEC architecture consists of three high level components depicted
in figure \ref{fig:hc_framework}:

@@

\subsection{EvoDraw}
As a case study, a IEA was developed using the 
ES-I platform \cite{garcia2013evospace}.
A brief description of the application is presented next, focusing
on the data elements that where ported to the graph model.

\subsection{Fitness Assignment}
\label{sec:assignment}
The ES-I % You should use always the same acronym, not start using it
         % here - JJ 
         % Done -Mario
 platform employs a collaborative methodology for performing fitness
 assignment. Several registered users assign a quality assessment to a single
phenotype and then an aggregated fitness value is calculated,
depending on the rating (from 1 to 5 stars) of each particular user,  % how do you define
                                % "taste"? - JJ
                                % Using actions
resulting in a many-to-many relationship between users and
phenotypes. 
 %scored by them? - JJ 
Other components or users of the systems can % formerly published methods? Other researchers? - JJ
% Corrected -Mario
query this user-phenotype relation to extract relevant
knowledge about the process and the population, for instance showing the
most popular phenotype, or the the user with more participation 
\cite{picbreeder}.
In order to do this, metadata about phenotypes % metadata better? - JJ Yes -Mario
must be permanently stored, even
if they are no longer in the active population.

\subsection{Collaboration}
\label{sec:col}
Users need to authenticate themselves to the system using their
Facebook account. In that sense, some users might not be interested
either because they do not want to give that information or simply because
they do not use that particular social network. Even as we might lose
some users that way, the additional information we obtain for scoring
phenotypes more than compensates for that. % Added this, check it
                                % out.  
                                % Ok! - Mario (and knowing how much more is future work)
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those phenotypes they like, or by taking phenotypes
from their friend's collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, a phenotype sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a rating % Why don't you use other
                                % thing instead of quality? We are not
                                % talking about quality, but about
                                % tastes or similar - JJ
                                % is rating Ok? 
 to the phenotype using
a five star rating system or,
additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special folder that stores those phenotypes a user likes and wishes
to save. After the user finishes interacting with the phenotype
on the Wall, he can choose to retrieve a new one from the population.
At the left hand side of \ref{fig:web}, the web page shows the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, users can browse the content of each collection and from
there share images through their social network.
When a user hovers the cursor over a phenotypes a pop-up pane shows how many users have
liked the phenotype. The pane also includes a link to the phenotype's 
details, the parents, genetic operators that created it, and genealogy
information. This makes the assignment of fitness through the rating
system a {\em social} activity, pursuing the objective of this work,
which is to increase engagement of the users.
\subsubsection{Graph Model}
